---
title: "Final"
author: "Vriz Gian Luca"
date: "2023-07-20"
output: pdf_document
---

#Libraries

```{r, message=FALSE}
library(exuber)
library(readr)
library(bsts)
library(dplyr)
library(car)
library(ModelMetrics)
library(hrbrthemes)
library(ggpmisc)
library(Rlibeemd)
library(xtable)
library(dtw)
library(lmtest)
library(Metrics)
library(tseries)
library(urca)
library(vars)
library(factoextra)
library(ggfortify)
library(modelr)
library(readxl)
library(epiDisplay)
library(blorr)
library(gridExtra)
library(oddsratio)
library(tictoc)
library(Rbeast)
library(psymonitor)
library(bbdetection)
library(devtools)
library(rbmi)
library(dplyr)
library(dqshiny)
library(ggplot2)
library(ghyp)
library(lubridate)
library(strucchange)
library(changepoint)
library(Rssa)
library(tidyr)
library(qcc)
library(corrplot)
library(zoo)
library(rugarch)
library(cpm)
library(visreg)
library(urca)
library(glmnet)
library(psych)
add_class <- function(x, ...) {
  class(x) <- append(c(...), class(x))
  x
}
```

#Bubbles

```{r}
#install_url('https://cran.r-project.org/src/contrib/Archive/psymonitor/psymonitor_0.0.2.tar.gz')
#install_url('https://cran.r-project.org/src/contrib/Archive/dqshiny/dqshiny_0.0.4.tar.gz')
add_class <- function(x, ...) {
  class(x) <- append(c(...), class(x))
  x
}
#DATA
#https://rdrr.io/rforge/rugarch/man/sp500ret.html
#https://rdrr.io/rforge/rugarch/man/dmbp.html
data("sp500ret")
data_garch = sp500ret*100
#GARCH
spec <- ugarchspec(variance.model = list(model = "eGARCH",garchOrder = c(1,1))
                   ,distribution.model="ghyp")
fit = ugarchfit(data = data_garch , spec = spec)

#residuals
garch_resid <- residuals(fit, standardize = FALSE)
# extract standardized residuals
garch_resid_std <- residuals(fit, standardize = TRUE)
# standardized residuals to be used in ugarchsim()
custom_dist = list(
  name = "sample",
  distfit = matrix(garch_resid_std,  ncol = 1)
)

m_ <- fit@model$maxOrder

garch_sim <- ugarchsim(fit, 
                       n.sim = length(garch_resid_std),
                       m.sim = 1, 
                       presigma = tail(fit@fit$sigma, m_),
                       prereturns = tail(sp500ret, m_),
                       # preresiduals ARE NOT standardized:
                       preresiduals = tail(garch_resid, m_),
                       startMethod = "sample",
                       # distfit in custom.dist ARE standardized
                       custom.dist = custom_dist)

# extract simulated series
sp_sim <- garch_sim@simulation$seriesSim
# test
sp_df <- cbind(sp500ret, sim = sp_sim) %>% 
  setNames(c("ret", "sim")) %>% 
  as_tibble(rownames = "date") %>% 
  mutate(date = as.Date(date),
         diff = sim - ret)

egarch<-function(n_sim) {
  
  garch_sim <- ugarchsim(fit, 
                         n.sim = n_sim,
                         m.sim = 1, 
                         presigma = tail(fit@fit$sigma, m_),
                         prereturns = tail(sp500ret, m_),
                         # preresiduals ARE NOT standardized:
                         preresiduals = tail(garch_resid, m_),
                         startMethod = "sample",
                         # distfit in custom.dist ARE standardized
                         #custom.dist = custom_dist
                         )
  values = as.numeric(garch_sim@simulation$seriesSim)
  return(values)
}
```

##Bubble collapse patterns

```{r}
bubble <- function(n, te = 0.4 * n, tf = te + 0.2 * n , tr = tf + 0.1*n,
                   c = 1, c1 = 1, c2 = 1, eta = 0.6, alpha = 0.6, beta = 0.5) {
  
  drift <- c*n^(-eta)
  delta <- 1 + c1 * n^(-alpha)
  gamma <- 1 - c2 * n^(-beta)
  y <- 100
  err <- egarch(n)*1.5
  
  for (t in 2:n) {
    if (t < te) {
      y[t] <- drift + y[t - 1] + err[t]
    } else if (t >= te & t <= tf) {
      y[t] <- delta * y[t - 1] + err[t]
    } else if (t > tf & t <= tr ) {
      y[t] <- gamma * y[t - 1] + err[t]
    } else {
      y[t] <- drift + y[t - 1] + err[t]
    }
  }
  y %>%
    add_class("sim") 
}


bubble_2 <- function(n,
                    te1 = 0.3 * n, tf1 = te1 + 0.1 * n , tr1 = tf1 + 0.1*n,
                    te2 = 0.8 * n, tf2 = te2 + 0.1 * n , tr2 = tf2 + 0.1*n,
                    c = 1, c1 = 1, c2 = 1, eta = 0.6, alpha = 0.6, beta = 0.7) {
  
  drift <- c*n^(-eta)
  delta <- 1 + c1 * n^(-alpha)
  gamma <- 1 - c2 * n^(-beta)
  y <- 100
  err <- egarch(n)
  
  for (t in 2:n) {
    if (t < te1) {
      y[t] <- drift + y[t - 1] + err[t] # normal
    } else if (t >= te1 & t <= tf1) {
      y[t] <- delta * y[t - 1] + err[t] # bubble1
    } else if (t > tf1 & t <= tr1 ) {
      y[t] <- gamma * y[t - 1] + err[t] # collapse 1
    }  else if (t > tr1 + 1 & t < te2) {
      y[t] <- drift + y[t - 1] + err[t] # normal 2
    }  else if (t >= te2 + 1 & t <= tf2) {
      y[t] <- delta * y[t - 1] + err[t] # bubble 2
    }  else if (t > tf2 + 1 & t <= tr2) {
      y[t] <- gamma * y[t - 1] + err[t] # collapse 2
    } else {
      y[t] <- drift + y[t - 1] + err[t] # normal 3
    }
  }
  y %>%
    add_class("sim")
}

set.seed(000)
time = 100
disturbing <- bubble(time, te = time*0.4, tf= time*0.6, tr = time*0.7, beta = 0.5, alpha = 0.6)
sudden <- bubble(time, time*0.5, tf= time*0.6, tr = time*0.66, beta = 0.3, alpha = 0.6, eta = 0.03)
smooth <- bubble(time, time*0.4, tf= time*0.6, tr = time*0.8, beta = 0.9, alpha = 0.6,eta=0.2)
```

##Simulations 

```{r}
autoplot(disturbing)
autoplot(smooth)
autoplot(sudden)
plot(1:99,diff(log(smooth)),type='l')
plot(1:99,diff(log(disturbing)),type='l')
plot(1:99,diff(log(sudden)),type='l')

plot(1:99,abs(diff(log(smooth))),type='l')
plot(1:99,abs(diff(log(disturbing))),type='l')
plot(1:99,abs(diff(log(sudden))),type='l')
```

```{r}
#TEST
sp500ret$data <- row.names(sp500ret)
sp500ret$data <- ymd(sp500ret$data)

series <- diff(log(disturbing))
              
test <- processStream(series,"Kolmogorov-Smirnov",ARL0=200, startup=17)
plot(1:length(series),series, type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = test$detectionTimes, col = 'blue')
abline(v = test$changePoints, lty = 2, col = 'red')
test

series <- diff(log(smooth))
              
test <- processStream(series,"Kolmogorov-Smirnov",ARL0=200, startup=17)
plot(1:length(series),series, type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = test$detectionTimes, col = 'blue')
abline(v = test$changePoints, lty = 2, col = 'red')
test

series <- diff(log(sudden))
              
test <- processStream(series,"Kolmogorov-Smirnov",ARL0=200, startup=17)
plot(1:length(series),series, type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = test$detectionTimes, col = 'blue')
abline(v = test$changePoints, lty = 2, col = 'red')
test

ewma <- numeric(length(series))
ewma[1] <- series[1]^2
lambda <- 0.94
for (i in 2:length(ewma)){
  ewma[i] <- lambda * ewma[i - 1] + (1 - lambda) * series[i]^2
}
plot(ewma, type = "l", xlab = "Observation", ylab = "", bty = "l")

abline(v = test$changePoints, lty = 2)

radf_sim <- radf(ts(disturbing))
autoplot(radf_sim)
datestamp(radf_sim)
```

##Disturbing

```{r}
#Distirubing(40,60,70)

matrix_empty = matrix(nrow = 100, ncol = 1000)

set.seed(002)
cyc = c(1:1000)
DIS <- data.frame(matrix_empty)
for (i in cyc){
  DIS[,i] <- bubble(time, te = time*0.4, tf= time*0.6, tr = time*0.7, beta = 0.5, alpha = 0.6)
}

results_DIS <- list()
ADF_DIS <- list()
for (i in cyc){
  tryCatch({ss <- diff(log(DIS[,i]))
  test <- processStream(ss,"Kolmogorov-Smirnov",ARL0=500, startup=17)
  results_DIS[[i]]<-test$changePoint
  remove(test)
  radf_sim <- radf(ts(DIS[,i]))
  d<-datestamp(radf_sim)
  ADF_DIS[[i]] <- c(d$series1$Start,d$series1$Peak,d$series1$End)
  remove(d)
  remove(radf_sim)
  print(i)
},error=function(e){cat('ERROR:',conditionMessage(e),'\n')})}
results_DIS
ADF_DIS
correct<-length(results_DIS[lengths(results_DIS) == 3])
correct_ADF<-length(ADF_DIS[lengths(ADF_DIS) == 3])
#914
#843
```

```{r}
matrix_test = matrix(ncol = 3, nrow = correct)
matrix_ADF = matrix(ncol = 3, nrow = correct_ADF)
for(i in 1:correct){
matrix_test[i,]<-c(results_DIS[lengths(results_DIS) == 3][[i]][1],results_DIS[lengths(results_DIS) == 3] [[i]][2],results_DIS[lengths(results_DIS) == 3][[i]][3])
}
for(i in 1:correct_ADF){
matrix_ADF[i,]<-c(as.numeric(substr(ADF_DIS[lengths(ADF_DIS) == 3][[i]][1], 2, 4)),as.numeric(substr(ADF_DIS[lengths(ADF_DIS) == 3][[i]][2], 2, 4)),as.numeric(substr(ADF_DIS[lengths(ADF_DIS) == 3][[i]][3], 2, 4)))
}

Metrics::rmse(matrix_test[,1],rep(40,correct))
Metrics::rmse(matrix_test[,2],rep(60,correct))
Metrics::rmse(matrix_test[,3],rep(70,correct))

Metrics::rmse(c(matrix_test[,1],matrix_test[,2],matrix_test[,3]),c(rep(40,correct),rep(60,correct),rep(70,correct)))

Metrics::rmse(matrix_ADF[,1],rep(40,correct_ADF))
Metrics::rmse(matrix_ADF[,2],rep(60,correct_ADF))
Metrics::rmse(matrix_ADF[,3],rep(70,correct_ADF))

Metrics::rmse(c(matrix_ADF[,1],matrix_ADF[,2],matrix_ADF[,3]),c(rep(40,correct_ADF),rep(60,correct_ADF),rep(70,correct_ADF)))
```

##Sudden

```{r}
#Sudden(50,60,66)
matrix_empty = matrix(nrow = 100, ncol = 1000)
set.seed(003)
SUD <- data.frame(matrix_empty)
for (i in cyc){
  SUD[,i] <- bubble(time, time*0.5, tf= time*0.6, tr = time*0.66, beta = 0.3, alpha = 0.6, eta = 0.03)
}

results_SUD <- list()
ADF_SUD <- list()
for (i in cyc){
  tryCatch({ss <- diff(log(SUD[,i]))
  test <- processStream(ss,"Kolmogorov-Smirnov",ARL0=500, startup=17)
  results_SUD[[i]]<-test$changePoints
  remove(test)
  radf_sim <- radf(ts(SUD[,i]))
  d<-datestamp(radf_sim)
  ADF_SUD[[i]] <- c(d$series1$Start,d$series1$Peak,d$series1$End)
  remove(d)
  remove(radf_sim)
  print(i)
  },error=function(e){cat('ERROR:',conditionMessage(e),'\n')})}
results_SUD
correct<-length(results_SUD[lengths(results_SUD) == 3])
correct_ADF<-length(ADF_SUD[lengths(ADF_SUD) == 3])
ADF_SUD
#846
#545
```

```{r}
matrix_test = matrix(ncol = 3, nrow = correct)
matrix_ADF = matrix(ncol = 3, nrow = correct_ADF)
for(i in 1:correct){
matrix_test[i,]<-c(results_SUD[lengths(results_SUD) == 3][[i]][1],results_SUD[lengths(results_SUD) == 3] [[i]][2],results_SUD[lengths(results_SUD) == 3] [[i]][3])
}
for(i in 1:correct_ADF){
matrix_ADF[i,]<-c(as.numeric(substr(ADF_SUD[lengths(ADF_SUD) == 3][[i]][1], 2, 4)),as.numeric(substr(ADF_SUD[lengths(ADF_SUD) == 3][[i]][2], 2, 4)),as.numeric(substr(ADF_SUD[lengths(ADF_SUD) == 3][[i]][3], 2, 4)))
}

Metrics::rmse(matrix_test[,1],rep(50,correct))
Metrics::rmse(matrix_test[,2],rep(60,correct))
Metrics::rmse(matrix_test[,3],rep(66,correct))

Metrics::rmse(c(matrix_test[,1],matrix_test[,2],matrix_test[,3]),c(rep(50,correct),rep(60,correct),rep(66,correct)))

Metrics::rmse(matrix_ADF[,1],rep(50,correct_ADF))
Metrics::rmse(matrix_ADF[,2],rep(60,correct_ADF))
Metrics::rmse(matrix_ADF[,3],rep(66,correct_ADF))

Metrics::rmse(c(matrix_ADF[,1],matrix_ADF[,2],matrix_ADF[,3]),c(rep(50,correct_ADF),rep(60,correct_ADF),rep(66,correct_ADF)))
```

##Smoothing

```{r}
matrix_empty = matrix(nrow = 100, ncol = 1000)
#Smooth(40,60,80)
set.seed(004)
SMO <- data.frame(matrix_empty)
for (i in cyc){
  SMO[,i] <- bubble(time, time*0.4, tf= time*0.6, tr = time*0.8, beta = 0.9, alpha = 0.6,eta=0.2)
}

results_SMO <- list()
ADF_SMO <- list()
for (i in cyc){
  tryCatch({ss <- diff(log(SMO[,i]))
  test <- processStream(ss,"Kolmogorov-Smirnov",ARL0=500, startup=17)
  results_SMO[[i]]<-test$changePoints
  remove(test)
  radf_sim <- radf(ts(SMO[,i]))
  d<-datestamp(radf_sim)
  ADF_SMO[[i]] <- c(d$series1$Start,d$series1$Peak,d$series1$End)
  remove(d)
  remove(radf_sim)
  print(i)
},error=function(e){cat('ERROR:',conditionMessage(e),'\n')})}
results_SMO
correct<-length(results_SMO[lengths(results_SMO) == 3])
correct_ADF<-length(ADF_SMO[lengths(ADF_SMO) == 3])
ADF_SMO
#885
#804
```

```{r}
matrix_test = matrix(ncol = 3, nrow = correct)
matrix_ADF = matrix(ncol = 3, nrow = correct_ADF)
for(i in 1:correct){
matrix_test[i,]<-c(results_SMO[lengths(results_SMO) == 3][[i]][1],results_SMO[lengths(results_SMO) == 3] [[i]][2],results_SMO[lengths(results_SMO) == 3] [[i]][3])
}
for(i in 1:correct_ADF){
matrix_ADF[i,]<-c(as.numeric(substr(ADF_SMO[lengths(ADF_SMO) == 3][[i]][1], 2, 4)),as.numeric(substr(ADF_SMO[lengths(ADF_SMO) == 3][[i]][2], 2, 4)),as.numeric(substr(ADF_SMO[lengths(ADF_SMO) == 3][[i]][3], 2, 4)))
}

Metrics::rmse(matrix_test[,1],rep(40,correct))
Metrics::rmse(matrix_test[,2],rep(60,correct))
Metrics::rmse(matrix_test[,3],rep(80,correct))

Metrics::rmse(c(matrix_test[,1],matrix_test[,2],matrix_test[,3]),c(rep(40,correct),rep(60,correct),rep(80,correct)))

Metrics::rmse(matrix_ADF[,1],rep(40,correct_ADF))
Metrics::rmse(matrix_ADF[,2],rep(60,correct_ADF))
Metrics::rmse(matrix_ADF[,3],rep(80,correct_ADF))

Metrics::rmse(c(matrix_ADF[,1],matrix_ADF[,2],matrix_ADF[,3]),c(rep(40,correct_ADF),rep(60,correct_ADF),rep(80,correct_ADF)))
```

#SP500

```{r}
#S&P500
SP500 <- as.data.frame(sp500d)
SP500$Date <- row.names(SP500)
SP500$Date <- ymd(SP500$Date)
rownames(SP500)<-NULL
SP500<-SP500[1:dim(SP500)[1],]
rownames(SP500)<-NULL
ggplot(SP500, aes(Date, SP500))+geom_line()
series <- ts(diff(log(SP500$SP500)),frequency=52)
ggplot(SP500[2:dim(SP500)[1],], aes(Date,series))+geom_line()+ggtitle('S&P500')+xlab('Time')+ylab('Absolute Returns')
```

##Daily

```{r}
tic('Time to run') 
test <- processStream(series,"Kolmogorov-Smirnov", ARL0=5000, startup=round(length(series)*17/100))
breack<-SP500$Date[test$changePoints]
plot(SP500$Date, log(SP500$SP500), type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = breack, lty = 2, col = 'red')
test

ewma <- numeric(length(series))
ewma[1] <- series[1]^2
lambda <- 0.94
for (i in 2:length(ewma)){
  ewma[i] <- lambda * ewma[i - 1] + (1 - lambda) * series[i]^2
}
plot(SP500$Date[2:dim(SP500)[1]], series, type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = breack, lty = 2, col = 'red')
toc()
```

##Monthly

```{r}
SP500 <- as.data.frame(sp500m)
SP500$Date <- row.names(SP500)
SP500$Date <- ymd(SP500$Date)
rownames(SP500)<-NULL
SP500<-SP500[1:dim(SP500)[1],]

rownames(SP500)<-NULL
ggplot(SP500, aes(Date, SP500))+geom_line()
#series <- ts(diff(emd(log(SP500$SP500),num_imfs = 2)[,2]),frequency = 52)
series <- ts(diff(log(SP500$SP500)),frequency=12)
ggplot(SP500[2:dim(SP500)[1],], aes(Date,series))+geom_line()+ggtitle('S&P500')+xlab('Time')+ylab('Absolute Returns')
```

```{r}
tic('Time to run') 
test <- processStream(series,"Kolmogorov-Smirnov", ARL0=500, startup=round(length(series)*1/100))
breack<-SP500$Date[test$changePoints]
plot(SP500$Date, log(SP500$SP500), type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = breack, lty = 2, col = 'red')
test

ewma <- numeric(length(series))
ewma[1] <- series[1]^2
lambda <- 0.94
for (i in 2:length(ewma)){
  ewma[i] <- lambda * ewma[i - 1] + (1 - lambda) * series[i]^2
}
plot(SP500$Date[2:dim(SP500)[1]], series, type = "l", xlab = "Observation", ylab = "", bty = "l")
abline(v = breack, lty = 2, col = 'red')
toc()
```

```{r}
tic('Time to run')
series<-ts(SP500$SP500,frequency=12,start=c(1950,1))
crit_values_sp<-radf_mc_cv(length(SP500$SP500), seed = 0005)
radf_sim <- radf(series)
datestamp(radf_sim)
autoplot(radf_sim)
toc()
```

#RENNIX

```{r}
renixx <- read.csv("Economic variables/Rennix_index.csv")
#DESCRIPTIVE STATISTICS
renixx$date<-ymd(renixx$date)
renixx<-renixx[order(renixx$date),]
rownames(renixx)<-NULL
renixx<-renixx[256:dim(renixx)[1],]
ggplot(renixx, aes(date, close))+geom_line()+ggtitle('Renixx index')+xlab('Time')+ylab('Absolute Returns')
```

##Daily

```{r}
renixx$date<-ymd(renixx$date)
df <- renixx %>% dplyr::select(1,3)
rownames(df) <- NULL
df$abs <- abs(c(0,diff(log(df$close))))

df$rate <- c(0,diff(log(df$close)))
ggplot(df, aes(date, abs))+geom_line()
```

```{r}
test <- processStream(ts(df$rate,freq=52,start=decimal_date(ymd("2005-1-1"))),"Kolmogorov-Smirnov", ARL0=5000, startup=round(dim(df)[1]*17/100)) #17% initial sample
breack<-df$date[test$changePoints]
allarm<-df$date[test$detectionTimes]
plot(df$date, df$close, type = "l", xlab = "Observation", ylab = "", bty = "l",main='Renixx')
abline(v = breack, lty = 2, col = 'red')
#abline(v = allarm, lty = 2, col = 'blue')
test
```

```{r}
breack
df$Date <- as.POSIXct(df$date, format = "%Y-%m-%d")
g<-ggplot(df, aes(x = Date, y=(close))) +
  geom_line(aes(color="RENIXX")) +
  theme_light()+ylab('Value')+xlab('Time')+ggtitle('Daily Data')+theme(plot.title = element_text(hjust = 0.9))+
   annotate("text", x = df$Date[730], y = df$close[730]+100, label = "Bubble Burst")+
  geom_segment(aes(x = df$Date[730], y = 0, xend = df$Date[730], yend =df$close[730]),linetype="dashed", color = "red")+
     annotate("text", x = df$Date[1364], y = df$close[1164]+100, label = "Bubble Decline")+
  geom_segment(aes(x = df$Date[1164], y = 0, xend = df$Date[1164], yend =df$close[1164]),linetype="dashed", color = "red")+
  annotate("text", x = df$Date[2016], y = df$close[2016]+250, label = "Bubble End")+
  geom_segment(aes(x = df$Date[2016], y = 0, xend = df$Date[2016], yend =df$close[2016]),linetype="dashed", color = "orange")+
  annotate("text", x = df$Date[2883], y = df$close[2883]+150, label = "Restore")+
  geom_segment(aes(x = df$Date[2883], y = 0, xend = df$Date[2883], yend =df$close[2883]),linetype="dashed", color = "green")+
      annotate("text", x = df$Date[4123], y = df$close[3823]-160, label = "Bubble Start")+
  geom_segment(aes(x = df$Date[3823], y = 0, xend = df$Date[3823], yend =df$close[3823]),linetype="dashed", color = "orange")+
annotate("rect", xmin = as.POSIXct(c(df$Date[1164],df$Date[2883])), xmax = as.POSIXct(c(df$Date[2016],df$Date[3823])), ymin = 0, ymax = Inf, alpha = .2,fill="orange")+
  annotate("rect", xmin = as.POSIXct(c(df$Date[1],df$Date[730],df$Date[3823])), xmax = as.POSIXct(c(df$Date[730],df$Date[1164],df$Date[dim(df)[1]])), ymin = 0, ymax = Inf, alpha = .2,fill="red")+
    annotate("rect", xmin = as.POSIXct(df$Date[2016]), xmax = as.POSIXct(df$Date[2883]), ymin = 0, ymax = Inf, alpha = .2,fill="green")+
  geom_point(data=df[c(730,1164),], aes(x = Date, y=(close)), colour="red", size=2)+
  geom_point(data=df[c(2016,3823),], aes(x = Date, y=(close)), colour="orange", size=2)+
  geom_point(data=df[c(2883),], aes(x = Date, y=(close)), colour="green", size=2)+
  scale_color_manual(name='Series',
                     breaks=c('RENIXX'),
                     values=c('RENIXX'='darkgreen'))
g
```

##Montlhy

```{r}
Month<- as.Date(cut(df$date, "month"))
renixx_m <- aggregate(close ~ Month, df, mean)
renixx_m$abs <- c(0,abs(diff(log(renixx_m$close))))

renixx_m$Renixx <- c(0,diff(log(renixx_m$close)))
renixx_m$St<-log(renixx_m$close)
ggplot(renixx_m, aes(Month, abs))+geom_line()+xlab('Time')+ylab('Absolute Returns')+ggtitle('Renixx index')
ggplot(renixx_m, aes(Month, close))+geom_line()+xlab('Time')+ylab('Value')+ggtitle('Renixx index')
```

```{r}
test <- processStream(ts(renixx_m$Renixx,freq=12,start=decimal_date(ymd("2005-1-1"))),"Kolmogorov-Smirnov", ARL0=500, startup=dim(renixx_m)[1]*17/100) #1 initial sample
breack<-renixx_m$Month[test$changePoints]
plot(renixx_m$Month, renixx_m$close, type = "l", xlab = "Observation", ylab = "", bty = "l",main='Renixx')
abline(v = breack, lty = 2, col = 'red')
test
```

```{r}
renixx_m$Range <- aggregate(high ~ Month, renixx, mean)[13:228,2]-aggregate(low ~ Month, renixx, mean)[13:228,2]
plot(renixx_m$Month[23:216], renixx_m$Range[23:216],type='l',main='Range price')
```

#Explosive behaviours

```{r}
set.seed(123)
series <- ts(renixx_m[,c(2,4)],frequency=12,start=c(2005,1))
results<-radf(series[,1])
autoplot(results)+labs(title = "Renixx BSADF")
d_adf<-datestamp(results)
d_adf
```

##Descriptive statistics

```{r}
#Oil WTI Futures
Oil <- read.csv("Economic variables/Crude Oil WTI Futures Historical Data_2005.csv")
Oil$Date <- mdy(Oil$Date)
Month <- as.Date(cut(Oil$Date, "month"))
Oil_m <- aggregate(Price ~ Month, Oil, mean)

Oil_m$Rate_O<-c(0,diff(log(Oil_m$Price)))
Oil_m$Rate_s<-scale(c(0,diff(Oil_m$Price)),scale=TRUE,center=TRUE)
rownames(Oil_m)<-NULL
ggplot(Oil_m, aes(Month, Price))+geom_line()+ggtitle('Oil price')
```

```{r}
#Geopolitical index 
data_gpr_export <- read_excel("Economic variables/data_gpr_daily_recent.xls")
geo_i <- data_gpr_export %>% dplyr::select(6,3)
colnames(geo_i)[1] <- "Date"
geo_i2 <- geo_i[geo_i$Date <= "2022-12-30" & geo_i$Date >= "2005-01-01", ]

Month <- as.Date(cut(geo_i2$Date, "month"))
geo_m <- aggregate(geo_i2$GPRD ~ Month,geo_i2, mean)
colnames(geo_m)[2] <- "GPRH"
geo_m$GPRH<-as.numeric(geo_m$GPRH)

geo_m$Rate_g <- c(0,diff(log(geo_m$GPRH)))
geo_m$St <- scale(c(0,diff(geo_m$GPRH)),scale=TRUE,center=TRUE)
ggplot(geo_m, aes(Month, GPRH))+geom_line()+ggtitle('Geopolitical index')
```

```{r}
#MSCI_WORLD
MSCI<-read_csv("Economic variables/MSCI World_2005.csv")
MSCI_2 <- MSCI %>% dplyr::select(6,2)
MSCI_2$Date <- format(as.Date(MSCI_2$Date,format='%m/%d/%y'), "%m/%d/%Y")
MSCI_2$Date <- mdy(MSCI_2$Date)
Month <- as.Date(cut(MSCI_2$Date, "month"))
MSCI_2$Close<-as.numeric(MSCI_2$Close)
MSCI_m <- aggregate(Close ~ Month, MSCI_2, mean)
colnames(MSCI_m)[2] <- "Price"
MSCI_m$Price<-as.numeric(MSCI_m$Price)

MSCI_m$Rate_ms <- c(0,diff(log(MSCI_m$Price)))
MSCI_m$Rate_s <- scale(c(0,diff(MSCI_m$Price)),scale=TRUE,center=TRUE)
ggplot(MSCI_m, aes(Month, Price))+geom_line()+ggtitle('Global MSCI index')
```

```{r}
renixx_m$Oil_ratio<-renixx_m$close/Oil_m$Price
plot(renixx_m$Month,renixx_m$Oil_ratio,type='l',xlab='Observations',ylab='Oil ratio')
renixx_m$MSCI_ratio<-renixx_m$close/MSCI_m$Price
plot(renixx_m$Month,renixx_m$MSCI_ratio,type='l',xlab='Observations',ylab='MSCI ratio')
```

```{r}
renixx_m$Oil_abs<-abs(c(0,diff(log(renixx_m$Oil_ratio))))
renixx_m$MSCI_abs<-abs(c(0,diff(log(renixx_m$MSCI_ratio))))
test <- processStream(renixx_m$Oil_abs,"Kolmogorov-Smirnov", ARL0=500, startup=dim(renixx_m)[1]*1/100) #1% initial sample
breack_2<-renixx_m$Month[test$changePoints]
plot(renixx_m$Month, renixx_m$Oil_ratio, type = "l", xlab = "Observation", ylab = "", bty = "l",main='Oil ratio')
abline(v = breack_2, lty = 2, col = 'red')

test <- processStream(renixx_m$MSCI_abs,"Kolmogorov-Smirnov", ARL0=500, startup=dim(renixx_m)[1]*1/100) #1% initial sample
breack_3<-renixx_m$Month[test$changePoints]
plot(renixx_m$Month, renixx_m$MSCI_ratio, type = "l", xlab = "Observation", ylab = "", bty = "l",main='MSCI ratio')
abline(v = breack_3, lty = 2, col = 'red')
```

```{r}
# create a plot object
renixx_m$Date <- as.POSIXct(renixx_m$Month, format = "%Y-%m-%d")
df_rect<-data.frame(xmin = c(as.POSIXct(d_adf$series1$Start)), xmax = c(as.POSIXct(d_adf$series1$End)), ymin = 0, ymax = Inf,Explosive=c("BSADF"))
g<-ggplot(renixx_m, aes(x = Date, y=(close/30))) +
  geom_line(aes(color="RENIXX")) +
  geom_line(aes(y = MSCI_ratio*10, color = "Oil ratio"), linetype="twodash") + 
  geom_line(aes(y = Oil_ratio/2, color="MSCI ratio"), linetype="twodash") +
  theme_light()+ylab('Value')+xlab('Time')+ggtitle('Empirical Analysis')+theme(plot.title = element_text(hjust = 0.5))+
  geom_segment(aes(x = renixx_m$Date[36], y = 0, xend = renixx_m$Date[36], yend =renixx_m$close[36]/30),linetype="dashed", color = "red")+
   annotate("text", x = renixx_m$Date[36], y = renixx_m$close[36]/30+5, label = "Bubble Burst")+
    geom_segment(aes(x = renixx_m$Date[95], y = 0, xend = renixx_m$Date[95], yend =renixx_m$close[95]/30),linetype="dashed", color = "red")+
  annotate("text", x = renixx_m$Date[95], y = renixx_m$close[95]/30+5, label = "Bubble End")+
      geom_segment(aes(x = renixx_m$Date[124], y = 0, xend = renixx_m$Date[124], yend =renixx_m$close[124]/30),linetype="dashed", color = "red")+
  annotate("text", x = renixx_m$Date[125], y = renixx_m$close[124]/30+5, label = "Restore")+
    geom_segment(aes(x = renixx_m$Date[178], y = 0, xend = renixx_m$Date[178], yend =renixx_m$close[178]/30),linetype="dashed", color = "red")+
  annotate("text", x = renixx_m$Date[170], y = renixx_m$close[178]/30+5, label = "Bubble Start")+
  geom_rect(data=df_rect,aes(xmin=xmin,ymin=ymin,xmax=xmax,ymax=ymax,fill=Explosive),
                    alpha=0.2,inherit.aes=FALSE)+
  scale_fill_manual(name='Test',values=c("azure4"))+
  geom_point(data=renixx_m[c(36,95,124,178), ], aes(x = Date, y=(close/30)), colour="red", size=2)+
  scale_color_manual(name='Time Series',
                     breaks=c('RENIXX', 'Oil ratio', 'MSCI ratio'),
                     values=c('RENIXX'='darkgreen', 'Oil ratio'='darkred', 'MSCI ratio'='steelblue'))

grid.arrange(g)
```

#Dataset

```{r}
#Policy risk and Financial risk
GEPUCURRENT <- read.csv("Economic variables/GEPUCURRENT.csv")
GEPUCURRENT$DATE<-ymd(GEPUCURRENT$DATE)
GEPUCURRENT<-GEPUCURRENT[97:312,]
rownames(GEPUCURRENT) <- NULL
GEPUCURRENT$Rate<-c(0,diff(log(GEPUCURRENT$GEPUCURRENT)))
GEPUCURRENT$Rate_s<-scale(c(0,diff(GEPUCURRENT$GEPUCURRENT)),scale=TRUE,center=TRUE)

Fs<-read.csv("Economic variables/Fs.csv")
Month<- as.Date(cut(ymd(Fs$Date), "month"))
Fs_m <- aggregate(OFR.FSI ~ Month, Fs, mean)
Fs_m<-Fs_m[61:276,]
Fs_m$Rate<-c(0,diff(Fs_m$OFR.FSI))
Fs_m$Rate_s<-scale(c(0,diff(Fs_m$OFR.FSI)),scale=TRUE,center=TRUE)
ggplot(Fs_m, aes(Month, OFR.FSI))+geom_line()+ggtitle('Financial Stability')
ggplot(GEPUCURRENT, aes(DATE, GEPUCURRENT))+geom_line()+ggtitle('Policy Uncertainty')
```

##Finance variables

```{r}
#Financial variables
Finance<-renixx_m[,c(1,4)]
colnames(Finance)[2] <- "Renixx"

Finance$Oil_p<-Oil_m$Rate_O

Finance$MSCI<-MSCI_m$Rate_ms

Finance$EPU<-GEPUCURRENT$Rate

Finance$OFR<-Fs_m$Rate

Finance$Geo_i<-geo_m$Rate_g

Finance_2<-renixx_m[,c(1,2)]
colnames(Finance_2)[2] <- "Renixx"
Finance_2$Renixx<-scale(Finance_2$Renixx,center=TRUE,scale=TRUE)
Finance_2$Oil_p<-scale(Oil_m$Price,center=TRUE,scale=TRUE)
Finance_2$Geo_i<-scale(geo_m$GPRH,center=TRUE,scale=TRUE)
Finance_2$MSCI<-scale(MSCI_m$Price,center=TRUE,scale=TRUE)
Finance_2$EPU<-scale(GEPUCURRENT$GEPUCURRENT,center=TRUE,scale=TRUE)
Finance_2$OFR<-scale(Fs_m$OFR.FSI,center=TRUE,scale=TRUE)


ggplot() + 
  geom_line(data = Finance_2, aes(x = Month, y = Renixx, color = "Rennix")) +
  geom_line(data = Finance_2, aes(x = Month, y = MSCI, color = "Global MSCI")) +
  geom_line(data = Finance_2, aes(x = Month, y = Geo_i, color = "Geopolitical index")) +
  geom_line(data = Finance_2, aes(x = Month, y = Oil_p, color = "Oil price")) +
  geom_line(data = Finance_2, aes(x = Month, y = EPU, color = "EPU")) +
  geom_line(data = Finance_2, aes(x = Month, y = OFR, color = "OFR")) +
  xlab('data_date') + ggtitle('Financial time series')+
  ylab('Monthly series')+ scale_color_manual(name='Series',
                     breaks=c('Rennix', 'Global MSCI','Geopolitical index','Oil price','EPU','OFR'),
                     values=c('Rennix'='green', 'Global MSCI'='red','Geopolitical index'='blue','Oil price'='brown','EPU'='orange','OFR'='grey'))
```

##Google trends

```{r}
#Text variables
green_e <- read.csv("Google Trends/Green_energy.csv", header=TRUE, comment.char="#")[13:228,]
green_e$Time<-ymd(green_e$Time)
green_e$Rate<-c(0,diff(log(green_e$Absolute.Google.Search.Volume)))
Climate<-as.data.frame(green_e[,1]) 
colnames(Climate)[1] <- "Date"
Climate$Green_e<-green_e[,4]

Tech_d <- read.csv("Google Trends/New Technology.csv", header=TRUE, comment.char="#")[13:228,]
Tech_d$Rate<-c(0,diff(log(Tech_d$Absolute.Google.Search.Volume)))
Climate$Tech<-Tech_d[,4]

Energy_i <- read.csv("Google Trends/Energy_index.csv", header=TRUE, comment.char="#")[13:228,]

Energy_i<-Energy_i[,c(1,2,3)]
Energy_i$Rate<-c(0,diff(log(Energy_i$Absolute.Google.Search.Volume)))
Climate$Energy_i<-Energy_i[,4]

global_w <- read.csv("Google Trends/Global_warming.csv", header=TRUE, comment.char="#")[13:228,]
global_w$Rate<-c(0,diff(log(global_w$Absolute.Google.Search.Volume)))
Climate$Warming<-global_w[,4]

Natural <- read.csv("Google Trends/Natural_disasters.csv", header=TRUE, comment.char="#")[13:228,]
Natural$Rate<-c(0,diff(log(Natural$Absolute.Google.Search.Volume)))
Climate$Natural<-Natural[,4]

Carbon_p <- read.csv("Google Trends/Carbon_price.csv", header=TRUE, comment.char="#")[13:228,]
Carbon_p$Rate<-c(0,diff(log(Carbon_p$Absolute.Google.Search.Volume)))
Climate$Carbon_p<-Carbon_p[,4]

Tax <- read.csv("Google Trends/Carbon_tax.csv", header=TRUE, comment.char="#")[13:228,]
Tax$Rate<-c(0,diff(log(Tax$Absolute.Google.Search.Volume)))
Climate$Tax<-Tax[,4]

Energy_s <- read.csv("Google Trends/Energy_shares.csv", header=TRUE, comment.char="#")[13:228,]
Energy_s$Rate<-c(0,diff(log(Energy_s$Absolute.Google.Search.Volume)))
Climate$Energy_s<-Energy_s[,4]

Climate_2<-as.data.frame(green_e[,1]) 
colnames(Climate_2)[1] <- "Date"

Climate_2$Green_e<-scale(as.numeric(green_e[,3]),center = TRUE,scale=TRUE)
Climate_2$Tech<-scale(as.numeric(Tech_d[,3]),center = TRUE,scale=TRUE)
Climate_2$Warming<-scale(as.numeric(global_w[,3]),center = TRUE,scale=TRUE)
Climate_2$Energy_i<-scale(as.numeric(Energy_i[,3]),center = TRUE,scale=TRUE)
Climate_2$Natural<-scale(as.numeric(Natural[,3]),center = TRUE,scale=TRUE)
Climate_2$Carbon_p<-scale(as.numeric(Carbon_p[,3]),center = TRUE,scale=TRUE)
Climate_2$Energy_s<-scale(as.numeric(Energy_s[,3]),center = TRUE,scale=TRUE)
Climate_2$Energy_T<-scale(as.numeric(Energy_s[,3]+Energy_i[,3])/2,center = TRUE,scale=TRUE)
Climate_2$Energy_s<-scale(as.numeric(Energy_s[,3]),center = TRUE,scale=TRUE)
Climate_2$Tax<-scale(as.numeric(Tax[,3]),center = TRUE,scale=TRUE)

ggplot() + 
  geom_line(data = Climate_2, aes(x = Date, y = Green_e, color = "Green Energy")) +
  geom_line(data = Climate_2, aes(x = Date, y = Energy_i, color = "Energy Index")) +
  geom_line(data = Climate_2, aes(x = Date, y = Warming, color = "Global Warming")) +
  geom_line(data = Climate_2, aes(x = Date, y = Natural, color = "Natural Disasters"))+ 
  geom_line(data = Climate_2, aes(x =Date, y = Carbon_p, color = "Carbon price")) +
 geom_line(data = Climate_2, aes(x =Date, y = Energy_s, color = "Energy shares")) +
  geom_line(data = Climate_2, aes(x =Date, y = Tax, color = "Carbon Tax")) +
  geom_line(data = Climate_2, aes(x =Date, y = Tech, color = "New Technology")) +
  xlab('data_date') + ggtitle('Text-time series')+
  ylab('Monthly series')+ scale_color_manual(name='Series',
                     breaks=c('Green Energy', 'Global Warming','Natural Disasters','Carbon price',"New Technology",'Energy Index','Green Bond','Carbon Emissions','Energy shares','Carbon Tax'),
                     values=c('Green Energy'='green','Global Warming'='red', 'Natural Disasters'='gold','Carbon price'='brown','New Technology'='purple','Energy Index'='blue','Carbon Emissions'='orange','Energy shares'='violet','Carbon Tax'='grey'))
```

```{r}
corrplot(cor(Climate[c(1:216),c(2:8)],method='kendall'), hclust.method='ward.D2',addrect=3,rect.col = 'gold',rect.lwd=4,order='hclust',diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')
corrplot(cor(Finance[c(1:216),(2:length(Finance[1,]))],method='kendall'), hclust.method='ward.D2',addrect=3,rect.col = 'gold',rect.lwd=4,order='hclust',diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')
```

```{r}
Finance$Merge<-format(as.Date(Finance$Month), "%Y-%m")
Climate$Merge<-format(as.Date(Climate$Date), "%Y-%m")

Data<-merge(Finance,Climate,by="Merge")
Data<- Data[c(2:216),-c(1,9)]
rownames(Data) <- NULL

corrplot(cor(Data[1:95,c(2:length(Data[1,]))],method='kendall'),diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')
corrplot(cor(Data[96:178,c(2:length(Data[1,]))],method='kendall'),diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')
corrplot(cor(Data[179:215,c(2:length(Data[1,]))],method='kendall'),diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')
corrplot(cor(Data[,c(2:length(Data[1,]))],method='kendall'),diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')
#"2016-03-01" "2019-10-01"
```

```{r}
Climate_2<-Climate_2[c('Date','Green_e','Tech','Warming','Natural','Carbon_p','Energy_i','Energy_s','Tax')]
Finance_2$Merge<-format(as.Date(Finance_2$Month), "%Y-%m")
Climate_2$Merge<-format(as.Date(Climate_2$Date), "%Y-%m")

Data_or<-merge(Finance_2,Climate_2,by="Merge")
Data_or<- Data_or[,-c(1,9)]
rownames(Data_or) <- NULL
```

#Preliminary statistic

```{r}
#Stationary assumption of log-returns

summary(ur.df(Climate[,"Carbon_p"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[,"Green_e"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[,"Tech"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[,"Warming"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[,"Natural"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[3:216,"Tax"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[,"Energy_s"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Climate[,"Energy_i"], type = "drift",lags = 12, selectlags = "AIC"))

summary(ur.df(Finance[,"Oil_p"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Finance[,"Renixx"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Finance[3:216,"MSCI"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Finance[,"EPU"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Finance[3:216,"OFR"], type = "drift",lags = 12, selectlags = "AIC"))
summary(ur.df(Finance[,"Geo_i"], type = "drift",lags = 12, selectlags = "AIC"))

kpss.test(Climate[,"Carbon_p"])
kpss.test(Climate[,"Tech"])
kpss.test(Climate[,"Natural"])
kpss.test(Climate[,"Green_e"])
kpss.test(Climate[,"Warming"])
kpss.test(Climate[,"Energy_s"])
kpss.test(Climate[,"Tax"])
kpss.test(Climate[,"Energy_i"])
kpss.test(Finance[,"Oil_p"])
kpss.test(Finance[,"Renixx"])
kpss.test(Finance[,"MSCI"])
kpss.test(Finance[,"Geo_i"])
kpss.test(Finance[,"EPU"])
kpss.test(Finance[,"OFR"])
bptest(lm(Finance[,"OFR"] ~ seq_along(Finance[,"OFR"])))
```

#Cluster analysis

```{r}
ratesa<-Data_or[1:95,]
ratesa$Month<-NULL;
ratesa<-ratesa
tr<-t(ratesa);
distMatrix <- dist(tr, method='DTW');
#hierarchical clustering
hc <- hclust(distMatrix,method="average")
plot(hc, main="Clustering by average")

ratesa<-Data_or[96:178,]
ratesa$Month<-NULL;
ratesa<-ratesa
tr<-t(ratesa);
distMatrix <- dist(tr, method='DTW');
#hierarchical clustering
hc <- hclust(distMatrix,method="average")
plot(hc, main="Clustering by average")

ratesa<-Data_or[179:216,]
ratesa$Month<-NULL;
ratesa<-ratesa
tr<-t(ratesa);
distMatrix <- dist(tr, method='DTW');
#hierarchical clustering
hc <- hclust(distMatrix,method="average")
plot(hc, main="Clustering by average")


ratesa<-Data_or[,]
ratesa$Month<-NULL;
ratesa<-ratesa
tr<-t(ratesa);
distMatrix <- dist(tr, method='DTW');
#hierarchical clustering
hc <- hclust(distMatrix,method="average")
plot(hc, main="Clustering by average")
```

#Factor analysis

```{r}
#https://rpubs.com/Pun_/Exploratory_factor_Analysis
#Promax Rotation. An oblique rotation, which allows factors to be correlated. This rotation can be calculated more quickly than a direct oblimin rotation, so it is useful for large datasets.

Data_FA<-renixx_m[,c(1,4)]
colnames(Data_FA)[2] <- "Renixx"
Data_FA$Renixxlag<-lag(Finance$Renixx)
Data_FA$Oil_lagp<-lag(Finance$Oil_p)
Data_FA$MSCIlag<-lag(Finance$MSCI)
Data_FA$EPUlag<-lag(Finance$EPU)
Data_FA$Geo_lagi<-lag(Finance$Geo_i)
Data_FA$OFRlag<-lag(Finance$OFR)

Data_FA$Warminglag<-lag(global_w[,4])
Data_FA$Naturallag<-lag(Natural[,4])
Data_FA$Carbon_plag<-lag(Carbon_p[,4])
Data_FA$Techlag<-lag(Tech_d[,4])
Data_FA$Energy_ilag<-lag(Energy_i[,4])
Data_FA$Green_elag<-lag(green_e[,4])
Data_FA$Taxlag<-lag(Tax[,4])
Data_FA$Energy_slag<-lag(Energy_s[,4])

Data_FA<-Data_FA[c(2:216),]
rownames(Data_FA)<-NULL

KMO(Data_FA[,c(3:length(Data_FA[1,]))])
Data_factor <-Data_FA[,c(3:length(Data_FA[1,]))]
Data_factor <-Data_factor[,KMO(Data_factor)$MSAi>0.7]
base::round(KMO(Data_factor)$MSA,2)

cortest.bartlett(Data_factor)

ev <- eigen(cor(Data_factor)) # get eigenvalues
ev$values

scree(Data_factor, pc=FALSE)

fa.parallel(Data_factor)

Nfacs <- 2 # This is for four factors. You can change this as needed.

fit <- fa(Data_factor, Nfacs, rotate = "none",fm="promax")

print(fit, digits=3, cutoff=0.3, sort=TRUE)

loads <- fit$loadings

fa.diagram(loads)

dim(fit$loadings)

fs <- factor.scores(Data_factor, fit)   
fs <- fs$scores
Data_FA$Factor_climate<-fs[,1]
Data_FA$Factor_finance<-fs[,2]

Data_FA$OFR<-Fs_m$Rate[2:216]
cor(Data_FA$Factor_climate[c(1:95)],Data_FA$OFR[c(1:95)],method='kendal')
cor(Data_FA$Factor_climate[c(96:178)],Data_FA$OFR[c(96:178)],method='kendal')
cor(Data_FA$Factor_climate[c(179:215)],Data_FA$OFR[c(179:215)],method='kendal')

cor(Data_FA$Factor_finance[c(1:95)],Data_FA$OFR[c(1:95)],method='kendal')
cor(Data_FA$Factor_finance[c(96:178)],Data_FA$OFR[c(96:178)],method='kendal')
cor(Data_FA$Factor_finance[c(179:215)],Data_FA$OFR[c(179:215)],method='kendal')

var_fact<-VAR(Data_FA[c(1:215),c('OFRlag','Factor_climate')],season = 12,p=1,type='const')
summary(var_fact)
set.seed(6)
causality(var_fact, cause = 'Factor_climate', boot=TRUE, boot.runs=1000)
causality(var_fact, cause = 'OFRlag', boot=TRUE, boot.runs=1000)
```

#PCA Bubble

```{r}
library(bestNormalize)
library(caret)
library(Rssa)
library(lattice)
library(fANCOVA)
library(stats)
library(HoRM)
library(WaveletComp)
library(parameters)
library(TSA)
library(fGarch)
library(fpp)
library(vars)
library(BVAR)
library(Metrics)
library(aTSA)
library(tsoutliers)
library(tsDyn)
library(goft)
library(bruceR)
library(plotmo)
library(grid)
library(factoextra)
library(ggfortify)
library(tvReg)
library(forecast)
```

```{r}
#PCA

Data_PCA<-Data_FA[c(1:95),c(3:16)]
data_PCA<-prcomp(Data_PCA,scale =TRUE,center = TRUE)

autoplot(data_PCA,loadings=TRUE,loadings.colour='Blue',loadings.label=TRUE,loadings.labal.size=3,data=Data_FA[c(1:95),],colour='Renixx')

stats<-summary(data_PCA)
importance<-stats$importance[3,]
importance[importance<=0.85]

fviz_contrib(data_PCA, choice = "var", axes = 1, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 2, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 5, top = 3)

x_results<-as.data.frame(data_PCA$x)
PCA_data<-as.data.frame(x_results$PC1)
names(PCA_data)[1]<-'PC1'
PCA_data$PC2<-x_results$PC2
PCA_data$PC3<-x_results$PC3
PCA_data$PC4<-x_results$PC4
PCA_data$PC6<-x_results$PC6
PCA_data$PC7<-x_results$PC7

cor<-matrix(ncol=1,nrow=length(PCA_data[,]))
for(i in 1:length(PCA_data[,])){
  cor[i]<-cor(Data_FA$Renixx[1:95],(PCA_data[,i]),method='kendal')
}

plot(abs(cor))
```

```{r}
Data_PCA<-Data_FA[c(96:178),c(3:16)]
data_PCA<-prcomp(Data_PCA,scale =TRUE,CENTER=TRUE)

autoplot(data_PCA,loadings=TRUE,loadings.colour='Blue',loadings.label=TRUE,loadings.labal.size=3,data=Data_FA[c(96:178),],colour='Renixx')

stats<-summary(data_PCA)
importance<-stats$importance[3,]
importance[importance<=0.85]

fviz_contrib(data_PCA, choice = "var", axes = 1, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 2, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 3, top = 3)

x_results<-as.data.frame(data_PCA$x)
PCA_data<-as.data.frame(x_results$PC1)
names(PCA_data)[1]<-'PC1'
PCA_data$PC2<-x_results$PC2
PCA_data$PC3<-x_results$PC3
PCA_data$PC4<-x_results$PC4
PCA_data$PC6<-x_results$PC6

cor<-matrix(ncol=1,nrow=length(PCA_data[,]))
for(i in 1:length(PCA_data[,])){
  cor[i]<-cor(Data_FA$Renixx[96:178],(PCA_data[,i]),method='kendal')
}

plot(abs(cor))
```

```{r}
Data_PCA<-Data_FA[c(179:215),c(3:16)]
data_PCA<-prcomp(Data_PCA,scale =TRUE,center=TRUE)

autoplot(data_PCA,loadings=TRUE,loadings.colour='Blue',loadings.label=TRUE,loadings.labal.size=3,data=Data_FA[c(179:215),],colour='Renixx')

stats<-summary(data_PCA)
importance<-stats$importance[3,]
importance[importance<=0.85]

fviz_contrib(data_PCA, choice = "var", axes = 1, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 2, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 4, top = 3)

x_results<-as.data.frame(data_PCA$x)
PCA_data<-as.data.frame(x_results$PC1)
names(PCA_data)[1]<-'PC1'
PCA_data$PC2<-x_results$PC2
PCA_data$PC3<-x_results$PC3
PCA_data$PC4<-x_results$PC4
PCA_data$PC6<-x_results$PC6

cor<-matrix(ncol=1,nrow=length(PCA_data[,]))
for(i in 1:length(PCA_data[,])){
  cor[i]<-cor(Data_FA$Renixx[179:215],(PCA_data[,i]),method='kendal')
}

plot(abs(cor))
```

##Explosive/No-explosive

```{r}
Data_FA$dummy <- as.factor(ifelse(Data_FA$Month >= "2007-07-01" & Data_FA$Month <= "2007-08-01"|Data_FA$Month >= "2007-10-01" & Data_FA$Month <= "2008-01-01"|Data_FA$Month >= "2015-04-01" & Data_FA$Month <= "2015-06-01"|Data_FA$Month >= "2019-12-01" & Data_FA$Month <= "2020-03-01"|Data_FA$Month >= "2020-07-01" & Data_FA$Month <= "2021-05-01"|Data_FA$Month >= "2021-11-01" & Data_FA$Month <= "2021-12-01", 1,0))

Data_exp<-Data_FA[Data_FA$dummy==1,c(4:length(Data_FA)-2)]
Data_noexp<-Data_FA[Data_FA$dummy==0,c(4:length(Data_FA)-2)]
```

```{r}
d_adf
data_PCA<-prcomp(Data_exp[,c(4:length(Data_exp)-2)],scale =TRUE)

autoplot(data_PCA,loadings=TRUE,loadings.colour='Blue',loadings.label=TRUE,loadings.labal.size=3,data=Data_exp,colour='Renixx')

stats<-summary(data_PCA)
importance<-stats$importance[3,]
importance[importance<=0.85]

fviz_contrib(data_PCA, choice = "var", axes = 1, top = 3)
fviz_contrib(data_PCA, choice = "var", axes = 2, top = 3)

x_results<-as.data.frame(data_PCA$x)
PCA_data<-as.data.frame(x_results$PC1)
names(PCA_data)[1]<-'PC1'
PCA_data$PC2<-x_results$PC2
PCA_data$PC3<-x_results$PC3
PCA_data$PC4<-x_results$PC4
PCA_data$PC5<-x_results$PC5

cor<-matrix(ncol=1,nrow=length(PCA_data[,]))
for(i in 1:length(PCA_data[,])){
  cor[i]<-cor(Data_exp$Renixx,(PCA_data[,i]),method='kendal')
}

plot(abs(cor))

corrplot(cor(Data_exp[,c(3:length(Data_exp)-2)],method='kendall'),diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')

corrplot(cor(Data_noexp[,c(3:length(Data_noexp)-2)],method='kendall'),diag=FALSE, method='square',bg='black',col=c('lightblue','green'),tl.col='black')


ratesa<-Data_or[Data_FA$dummy==1,]
ratesa$Month<-NULL;
ratesa<-ratesa
tr<-t(ratesa);
distMatrix <- dist(tr, method='DTW');
#hierarchical clustering
hc <- hclust(distMatrix,method="average")
plot(hc, main="Clustering by average")


ratesa<-Data_or[Data_FA$dummy==0,]
ratesa$Month<-NULL;
ratesa<-ratesa
tr<-t(ratesa);
distMatrix <- dist(tr, method='DTW');
#hierarchical clustering
hc <- hclust(distMatrix,method="average")
plot(hc, main="Clustering by average")

cor(Data_exp$Factor_climate,Data_exp$Renixx,method='kendal')
cor(Data_exp$Factor_finance,Data_exp$Renixx,method='kendal')
cor(Data_exp$Factor_climate,Data_exp$OFRlag,method='kendal')
cor(Data_exp$Factor_finance,Data_exp$OFRlag,method='kendal')

tvLM<-tvLM(Renixx ~ 0 +Renixxlag+Energy_ilag+Energy_slag+Carbon_plag+Green_elag+Taxlag+MSCIlag+Warminglag+Naturallag+Techlag, data = Data_FA, bw = 0.5)
plot(Data_FA$Month, tvLM$coefficients[,1],xlab='Time',ylab='Renixx previous values correlation')
plot(Data_FA$Month, tvLM$coefficients[,2],xlab='Time',ylab='Energy index correlation')
```

#Market dynamics

##Bubbles dynamics

```{r}
#Dummy
breack
breack_2
breack_3
Data_2<-renixx_m[c(2:216),c(1,4)]
colnames(Data_2)[2] <- "Renixx"

Data_2$dummy_Renixx  <- cut(Data_2$Month,
              breaks= as.Date(c("2005-01-01","2012-12-01","2019-10-01",'2023-01-01')),
              labels=c('Clean-Tech', 'No-bubble', 'Climate'))

Data_2$dummy_Renixx<-relevel(Data_2$dummy_Renixx, ref = 'No-bubble')

Data_2$Energy_ilag<-lag(Data_or$Energy_i)[2:216]
Data_2$Warminglag<-lag(Data_or$Warming)[2:216]
Data_2$Naturallag<-lag(Data_or$Natural)[2:216]
Data_2$Green_elag<-lag(Data_or$Green_e)[2:216]
Data_2$Carbon_plag<-lag(Data_or$Carbon_p)[2:216]
Data_2$Techlag<-lag(Data_or$Tech)[2:216]
Data_2$Energy_slag<-lag(Data_or$Energy_s)[2:216]
Data_2$Taxlag<-lag(Data_or$Tax)[2:216]

Data_2<-Data_2[,-c(1,2)]

rownames(Data_2) <- NULL

str(Data_2)
```

```{r}
p1 <- ggplot(Data_2, aes(x=Green_elag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx))+ labs(fill='Bubble')+
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Green Energy') + ylab('Density')
p2 <- ggplot(Data_2, aes(x=Naturallag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4)+
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Natural Disasters')+ labs(fill='Bubble') + ylab('Density')
p3 <- ggplot(Data_2, aes(x=Carbon_plag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) + xlab('Carbon Price')+
      guides(color = "none")+ geom_density(alpha=0.4) + labs(fill='Bubble') + ylab('Density')
p4 <- ggplot(Data_2, aes(x=Energy_ilag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Energy Index')+ labs(fill='Bubble') + ylab('Density')
p5 <- ggplot(Data_2, aes(x=Warminglag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Global Warming')+ labs(fill='Bubble')+ ylab('Density')
p6 <- ggplot(Data_2, aes(x=Techlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('New Technology')+ labs(fill='Bubble')+ ylab('Density')
p7 <- ggplot(Data_2, aes(x=Energy_slag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('Energy Shares')+ labs(fill='Bubble')+ ylab('Density')
p8 <- ggplot(Data_2, aes(x=Taxlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('Carbon Tax')+ labs(fill='Bubble')+ ylab('Density')
grid.arrange(p1, p2, p3, p4, p5, p6,p7,p8, ncol=4,top = textGrob("RENIXX Index",gp=gpar(fontsize=20,font=3)))
```

```{r}
smp_size <- floor(0.7 * nrow(Data_2))
Data_multi<-Data_2
Data_multi$Time<-c(1:215)
##set the seed to make your partition reproducible
set.seed(1239)
train_ind <- sample(seq_len(nrow(Data_multi)), size = smp_size)

train.data <- Data_multi[train_ind, ]
vali.data <- Data_multi[-train_ind, ] 
#Dumy code categorical predictor variables
x <- model.matrix(dummy_Renixx ~., train.data)[,-1]
y<-train.data$dummy_Renixx
```

```{r}
#Find the best lambda using cross-validation
set.seed(123) 
cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "multinomial")
#Fit the final model on the training data
model <- glmnet(x, y, alpha = 1, family = "multinomial",
                lambda = cv.lasso$lambda.min)
#Display regression coefficients
coef(model)
# Display regression coefficients
plot_glmnet(cv.lasso$glmnet.fit, label=TRUE,nresponse = 3)
plot_glmnet(cv.lasso$glmnet.fit, label=TRUE,nresponse = 2)
plot_glmnet(cv.lasso$glmnet.fit, label=TRUE,nresponse = 1)
```

```{r}
set.seed(123)
cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "multinomial")
plot(cv.lasso)
```

```{r}
cv.lasso$lambda.min
coef(cv.lasso, cv.lasso$lambda.min)
```

```{r}
lasso.model <- glmnet(x, y, alpha = 1, family = "multinomial",
                      lambda = cv.lasso$lambda.min)
# Make prediction on test data
x.test <- model.matrix(dummy_Renixx ~., vali.data)[,-1]
probabilities <- lasso.model %>% predict(newx = x.test)
predicted.classes <- probabilities

for(i in 1:length(probabilities[,1,])){
  for(j in 1:length(probabilities[1,,])){
predicted.classes[i,j,] <- ifelse(probabilities[i,j,]==max(probabilities[i,,]), 1, 0)
}
}

predicted.classes

# 5 errors on 65
```

##Elastic-net regularized logistic regression for explosive moments

```{r}
df <- renixx %>% dplyr::select(1,5)
Month<- as.Date(cut(df$date, "month"))
renixx_m$low <- aggregate(low ~ Month, df, mean)[13:228,2]

df <- renixx %>% dplyr::select(1,4)
Month<- as.Date(cut(df$date, "month"))
renixx_m$high <- aggregate(high ~ Month, df, mean)[13:228,2]
```

```{r}
Data_2$dummy_Renixx<- as.factor(ifelse(Data_FA$Month >= "2007-07-01" & Data_FA$Month <= "2007-08-01"|Data_FA$Month >= "2007-10-01" & Data_FA$Month <= "2008-01-01"|Data_FA$Month >= "2015-04-01" & Data_FA$Month <= "2015-06-01"|Data_FA$Month >= "2019-12-01" & Data_FA$Month <= "2020-03-01"|Data_FA$Month >= "2020-07-01" & Data_FA$Month <= "2021-05-01"|Data_FA$Month >= "2021-11-01" & Data_FA$Month <= "2021-12-01", 'Explosive','No-Explosive'))
Data_2$dummy_Renixx
Data_2$dummy_Renixx<-relevel(Data_2$dummy_Renixx, ref = 'No-Explosive')

Data_2$diff_Enii<-lag(scale((Data_or$Renixx),center=TRUE,scale=TRUE))[2:216]
Data_2$MSCIlag<-lag(MSCI_m$Price)[2:216]
Data_2$Oillag<-lag(Oil_m$Price)[2:216]
Data_2$EPUlag<-lag(GEPUCURRENT$GEPUCURRENT)[2:216]
Data_2$OFRlag<-lag(Fs_m$OFR.FSI)[2:216]

Data_2$Oillag<-scale(c(0,diff(Data_2$Oillag)),center = TRUE,scale=TRUE)
Data_2$EPUlag<-scale(c(0,diff(Data_2$EPUlag)),center = TRUE,scale=TRUE)
Data_2$MSCIlag<-scale(c(0,diff(Data_2$MSCIlag)),center = TRUE,scale=TRUE)
Data_2$OFRlag<-scale(c(0,diff(Data_2$OFRlag)),center = TRUE,scale=TRUE)
```

```{r}
p1 <- ggplot(Data_2, aes(x=Green_elag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx))+ labs(fill='Green bubble')+
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Green Energy')
p2 <- ggplot(Data_2, aes(x=Naturallag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4)+
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Natural disaster')+ labs(fill='Green bubble') 
p3 <- ggplot(Data_2, aes(x=Carbon_plag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) + xlab('Carbon price')+
      guides(color = "none")+ geom_density(alpha=0.4) + labs(fill='Green bubble') 
p4 <- ggplot(Data_2, aes(x=Energy_ilag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Energy Index')+ labs(fill='Green bubble') 
p5 <- ggplot(Data_2, aes(x=Warminglag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('Global Warming')+ labs(fill='Green bubble')
p6 <- ggplot(Data_2, aes(x=Techlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('New Technology')+ labs(fill='Green bubble')
p7 <- ggplot(Data_2, aes(x=Energy_slag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('Energy Shares')+ labs(fill='Green bubble')
p8 <- ggplot(Data_2, aes(x=Taxlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('Carbon Tax')+ labs(fill='Green bubble')
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol=2,top = textGrob("Renixx Global Index Explosive Evets",gp=gpar(fontsize=20,font=3)))
```

```{r}
p9 <- ggplot(Data_2, aes(x=OFRlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('OFR')+ labs(fill='Green bubble')
p10 <- ggplot(Data_2, aes(x=MSCIlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
      guides(color = "none")+ geom_density(alpha=0.4) + xlab('MSCI')+ labs(fill='Green bubble')
p11 <- ggplot(Data_2, aes(x=EPUlag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('EPU')+ labs(fill='Green bubble')
p12 <- ggplot(Data_2, aes(x=Oillag, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('Oil price')+ labs(fill='Green bubble')

p13 <- ggplot(Data_2, aes(x=diff_Enii, group=dummy_Renixx, color=dummy_Renixx, fill=dummy_Renixx)) + geom_density(alpha=0.4) +
     guides(color = "none")+ geom_density(alpha=0.4) + xlab('Dfference energy index')+ labs(fill='Green #bubble')

grid.arrange(p9, p10, p11, p12, ncol=2,top = textGrob("Renixx Global Index Explosive Evets",gp=gpar(fontsize=20,font=3)))
```

```{r}
Data_2$dummy_Renixx<- as.factor(ifelse(Data_FA$Month >= "2007-07-01" & Data_FA$Month <= "2007-08-01"|Data_FA$Month >= "2007-10-01" & Data_FA$Month <= "2008-01-01"|Data_FA$Month >= "2015-04-01" & Data_FA$Month <= "2015-06-01"|Data_FA$Month >= "2019-12-01" & Data_FA$Month <= "2020-03-01"|Data_FA$Month >= "2020-07-01" & Data_FA$Month <= "2021-05-01"|Data_FA$Month >= "2021-11-01" & Data_FA$Month <= "2021-12-01", 1,0))

Data_2$Time<-scale(c(1:215),center=TRUE,scale=TRUE)
train.data  <- Data_2[c(1:190),]
vali.data <- Data_2[c(191:215),]
#Dumy code categorical predictor variables
x <- model.matrix(dummy_Renixx ~., train.data)[,-1]
#Convert the outcome (class) to a numerical variable
y <- train.data$dummy_Renixx
```

```{r}
set.seed(123)
seeds <- vector(mode = "list", length = 65)
for(i in 1:64) seeds[[i]] <- sample.int(1000, 15)
seeds[[65]] <- sample.int(1000, 1)
tuneLength.num <- 15
myTimeControl <- trainControl(method = "timeslice",
                              initialWindow = 135,
                              horizon = 20,
                              fixedWindow = FALSE,
                              allowParallel = FALSE,
                              seeds = seeds)

glmnet.mod <- train(dummy_Renixx ~.,
                    data = train.data,
                    method = "glmnet",
                    family = "binomial",
                    trControl = c(myTimeControl),
                    tuneLength=tuneLength.num) 
glmnet.mod
```

```{r}
#Fit the final model on the training data
model <- glmnet(x, y, alpha = glmnet.mod$bestTune$alpha, family = "binomial",
                lambda =glmnet.mod$bestTune$lambda)
model_plot <- glmnet(x, y, alpha = glmnet.mod$bestTune$alpha, family = "binomial")
#Display regression coefficients
coef(model)
#Make predictions on the test data
x.test <- model.matrix(dummy_Renixx ~., vali.data)[,-1]
probabilities <- model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, 1, 0)
#Model accuracy
observed.classes <- vali.data$dummy_Renixx
mean(predicted.classes == observed.classes)
confusionMatrix(as.factor(predicted.classes[,1]), observed.classes)

plot(model_plot, xvar = "lambda", label=T)
lbs_fun <- function(fit, offset_x=1, ...) {
L <- length(fit$lambda)
x <- log(fit$lambda[L])+ offset_x
y <- fit$beta[, L]
labs <- names(y)
text(x, y, labels=labs, ...)
}
```

##Logit model

```{r}
model_f<-glm(dummy_Renixx ~., family = binomial(link = "logit"), data = Data_2[1:190,c(1:14)])
model_0<-glm(dummy_Renixx ~1, family = binomial(link = "logit"), data = Data_2[1:190,])
model_2<-glm(dummy_Renixx ~MSCIlag+diff_Enii+Energy_ilag+Techlag, family = binomial(link = "logit"), data = Data_2[1:190,])
summary(model_2)
model_2$deviance
-2*logLik(model_2)

vif(model_2)
lrtest (model_0, model_2)
lrtest (model_2, model_f)
blr_linktest(model_2)

y_hat2 <- predict(model_2, type='response')
mean((y_hat2 >= 0.5 & Data_2[,]$dummy_Renixx==0) | (y_hat2 < 0.5 & Data_2[,]$dummy_Renixx==1))
1 - max(prop.table(table(Data_2[,]$dummy_Renixx)))

list(model = pscl::pR2(model_2)["McFadden"])

model1_data <- augment(model_2) %>% mutate(index = 1:n())

ggplot(model1_data, aes(index, .std.resid, color = dummy_Renixx)) + 
  geom_point(alpha = .5) +
  geom_ref_line(h = 3)

or_glm(data = Data_2[1:190,], model = model_2, incr = c(Naturallag=0.1,Green_elag=0.1,Energy_ilag=0.1,Techlag=0.1,Carbon_plag=0.1,Warminglag=0.1,Energy_slag=0.1,Taxlag=0.1,MSCIlag=0.1,diff_Enii=0.1,Time=0.1,OFRlag=0.1))
```

```{r}
predictions <- predict(model_2, type = "response", newdata=Data_2[191:215,])
predictions

threshold <- 0.5
binary_predictions <- ifelse(predictions > threshold, 1, 0)
confusionMatrix(as.factor(binary_predictions), observed.classes)
```

##VAR

```{r}
#granger causality
Finance$Merge<-format(as.Date(Finance$Month), "%Y-%m")
Climate$Merge<-format(as.Date(Climate$Date), "%Y-%m")

Data_3<-merge(Finance,Climate,by="Merge")
Data_3<- Data_3[c(1:216),-c(1,9)]
rownames(Data_3) <- NULL

grangertest(MSCI ~ Renixx, order =12, data = Data_3)
grangertest(MSCI ~ Oil_p, order =12, data = Data_3)#Granger
grangertest(MSCI ~ Geo_i, order =12, data = Data_3)
grangertest(MSCI ~ Energy_i, order =12, data = Data_3)
grangertest(MSCI ~ Green_e, order =12, data = Data_3)
grangertest(MSCI ~ Tech, order =12, data = Data_3)
grangertest(MSCI ~ Warming, order =12, data = Data_3)
grangertest(MSCI ~ Natural, order =12, data = Data_3)
grangertest(MSCI ~ Carbon_p, order =12, data = Data_3)
grangertest(MSCI ~ Tax, order =12, data = Data_3)
grangertest(MSCI ~ Energy_s, order =12, data = Data_3)
grangertest(MSCI ~ EPU, order =12, data = Data_3)
grangertest(MSCI ~ OFR, order =12, data = Data_3)

grangertest(Renixx ~ MSCI, order = 12, data = Data_3)
grangertest(Renixx ~ Oil_p, order = 12, data = Data_3)#Granger
grangertest(Renixx ~ Geo_i, order = 12, data = Data_3)
grangertest(Renixx ~ EPU, order = 12, data = Data_3)#Granger
grangertest(Renixx ~ OFR, order = 12, data = Data_3)
grangertest(Renixx ~ Energy_i, order = 12, data = Data_3)
grangertest(Renixx ~ Green_e, order = 12, data = Data_3)
grangertest(Renixx ~ Tech, order = 12, data = Data_3)
grangertest(Renixx ~ Warming, order = 12, data = Data_3)
grangertest(Renixx ~ Natural, order = 12, data = Data_3)
grangertest(Renixx ~ Carbon_p, order = 12, data = Data_3)
grangertest(Renixx ~ Energy_s, order = 12, data = Data_3)
grangertest(Renixx ~ Tax, order = 12, data = Data_3)#Granger

grangertest(Oil_p ~ MSCI, order = 12, data = Data_3)#Granger
grangertest(Oil_p ~ Natural, order = 12, data = Data_3)
grangertest(Oil_p ~ Geo_i, order = 12, data = Data_3)
grangertest(Oil_p ~ Renixx, order = 12, data = Data_3)
grangertest(Oil_p ~ Green_e, order = 12, data = Data_3)
grangertest(Oil_p ~ Tech, order = 12, data = Data_3)
grangertest(Oil_p ~ Warming, order = 12, data = Data_3)
grangertest(Oil_p ~ Energy_i, order = 12, data = Data_3)
grangertest(Oil_p ~ Carbon_p, order = 12, data = Data_3)
grangertest(Oil_p ~ EPU, order = 12, data = Data_3)
grangertest(Oil_p ~ OFR, order = 12, data = Data_3)#Granger

grangertest(OFR ~ MSCI, order = 12, data = Data_3)#Granger
grangertest(OFR ~ Oil_p, order = 12, data = Data_3)#Granger
grangertest(OFR ~ Geo_i, order = 12, data = Data_3)
grangertest(OFR ~ Renixx, order = 12, data = Data_3)#Granger
grangertest(OFR ~ Green_e, order = 12, data = Data_3)
grangertest(OFR ~ Tech, order = 12, data = Data_3)
grangertest(OFR ~ Warming, order = 12, data = Data_3)
grangertest(OFR ~ Natural, order = 12, data = Data_3)
grangertest(OFR ~ Carbon_p, order = 12, data = Data_3)
grangertest(OFR ~ Tax, order = 12, data = Data_3)#Granger
grangertest(OFR ~ EPU, order = 12, data = Data_3)
grangertest(OFR ~ Energy_i, order = 12, data = Data_3)
grangertest(OFR ~ Energy_s, order = 12, data = Data_3)
```

```{r}
Data_or$Energy_n<-scale(Energy_i$Absolute.Google.Search.Volume+Energy_s$Absolute.Google.Search.Volume,center=TRUE,scale=TRUE)
ggplot() + 
     geom_line(data = Data_or, aes(x = Month, y = Renixx, color = "Renixx")) +
     geom_line(data = Data_or, aes(x = Month, y = Energy_i, color = 'Energy Index'))+
  geom_line(data = Data_or, aes(x = Month, y = Energy_s, color = 'Energy Stock'))+
  xlab('Date') + ggtitle('Series')+
  ylab('Monthly series')+ scale_color_manual(name='Series',
                     breaks=c('Renixx', 'Energy Index', 'Energy Stock'),
                     values=c('Renixx'='red', 'Energy Index'='blue','Energy Stock'='darkgreen'))

cor(Data_or$Renixx,Data_or$Energy_i)
cor(Data_or$Energy_i[121:216],Data_or$Renixx[121:216])
cor(Data_or$Energy_s[121:216],Data_or$Renixx[121:216])
cor(Data_or$Oil_p[121:216],Data_or$Renixx[121:216])
```

#Forecasting

```{r}
Data_VECM<-as.data.frame(log(renixx_m$close))
colnames(Data_VECM)[1]<-'Renixx'

Data_VECM$Energy_i<-log(Energy_i$Absolute.Google.Search.Volume)

Data_VECM$Energy_s<-log(Energy_s$Absolute.Google.Search.Volume)

Data_VECM$Oil_p<-log(Oil_m$Price)
```

```{r}
grangertest(Renixx ~ Energy_i, order = 12, data = Data_3[121:216,])#Granger
grangertest(Renixx ~ Energy_s, order = 12, data = Data_3[121:216,])#Granger
grangertest(Renixx ~ Oil_p, order = 12, data = Data_3[121:216,])#Granger
#Granger
grangertest(Energy_s ~ Renixx, order = 12, data = Data_3[121:216,])
grangertest(Energy_i ~ Renixx, order = 12, data = Data_3[121:216,])
grangertest(Oil_p ~ Renixx, order = 12, data = Data_3[121:216,])

grangertest(Renixx ~ MSCI, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Tax, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Oil_p, order = 12, data = Data_3[121:216,])#Granger
grangertest(Renixx ~ Geo_i, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Green_e, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Tech, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Warming, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Natural, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Carbon_p, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ EPU, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ OFR, order = 12, data = Data_3[121:216,])
grangertest(Renixx ~ Energy_i, order = 12, data = Data_3[121:216,])#Granger
grangertest(Renixx ~ Energy_s, order = 12, data = Data_3[121:216,])#Granger

grangertest(Renixx ~ Energy_i, order = 12, data = Data_3[121:216,])#Granger
grangertest(Renixx ~ Energy_s, order = 12, data = Data_3[121:216,])#Granger
grangertest(Renixx ~ Oil_p, order = 12, data = Data_3[121:216,])#Granger

grangertest(Energy_s ~ Renixx, order = 12, data = Data_3[121:216,])
grangertest(Energy_i ~ Renixx, order = 12, data = Data_3[121:216,])
grangertest(Oil_p ~ Renixx, order = 12, data = Data_3[121:216,])

grangertest(Energy_s ~ Energy_i, order = 12, data = Data_3[121:216,])
grangertest(Energy_s ~ Oil_p, order = 12, data = Data_3[121:216,])
grangertest(Oil_p ~ Energy_i, order = 12, data = Data_3[121:216,])
grangertest(Oil_p ~ Energy_s, order = 12, data = Data_3[121:216,])
grangertest(Energy_i ~ Energy_s, order = 12, data = Data_3[121:216,])
grangertest(Energy_i ~ Oil_p, order = 12, data = Data_3[121:216,])
```

##One step ahead

```{r}
Energy_ipred <- read.csv("Predictions/Energy_spred.csv", header=TRUE, comment.char="#")
Energy_spred <- read.csv("Predictions/Energy_ipred.csv", header=TRUE, comment.char="#")
Oil_pred <- read.csv("Predictions/Oil_pred.csv", header=TRUE, comment.char="#")
Oil_pred$Date <- mdy(Oil_pred$Date)
Month <- as.Date(cut(Oil_pred$Date, "month"))
Oil_predm <- aggregate(Price ~ Month, Oil_pred, mean)
Renixx_pred <- read.csv("Predictions/Renixx_pred.csv", header=TRUE, comment.char="#")
Renixx_pred$date <- ymd(Renixx_pred$date)
Month <- as.Date(cut(Renixx_pred$date, "month"))
Renixx_predm <- aggregate(close ~ Month, Renixx_pred, mean)
data_pred0<-data.frame(log(Renixx_predm$close),log(Energy_ipred[229:239,3]),log(Oil_predm$Price),log(Energy_spred[229:239,3]))
colnames(data_pred0)<-c('Renixx','Energy_i','Oil_p',"Energy_s")
data_pred<-rbind(Data_VECM[c(121:216),c('Renixx','Energy_i','Oil_p',"Energy_s")],data_pred0)
rownames(data_pred)<-NULL
```

```{r}
iteration<-11
models<-6
sample<-length(data_pred[,1])
predictions_uni<-as.data.frame(matrix(nrow=iteration,ncol=models))
colnames(predictions_uni)<-c('ARIMA','ARIMAX','ARIMAX2','VAR','BSTS','NNETAR')
up_uni<-as.data.frame(matrix(nrow=iteration,ncol=models))
colnames(up_uni)<-c('ARIMA','ARIMAX','ARIMAX2','VAR','BSTS','NNETAR')
lower_uni<-as.data.frame(matrix(nrow=iteration,ncol=models))
colnames(lower_uni)<-c(c('ARIMA','ARIMAX','ARIMAX2','VAR','BSTS','NNETAR'))
```

```{r}
#Arima
all_series<-ts(data_pred,start=c(2015,1),frequency=12)
for(i in 96:(sample-1)){
arima<-auto.arima(all_series[1:i,c('Renixx')],seasonal = TRUE)
pred_arima<-forecast::forecast(arima,h=1)
predictions_uni[(i-96+1),1]<-pred_arima$mean
lower_uni[(i-96+1),1]<-pred_arima$lower[,2]
up_uni[(i-96+1),1]<-pred_arima$upper[,2]
}
```

```{r}
#Arimax with oil
data_pred2<-cbind(data_pred$Renixx,lag(data_pred[,2:4]))[2:length(data_pred[,1]),]
rownames(data_pred2)<-NULL
colnames(data_pred2)<-c('Renixx','Energy_i','Oil_p',"Energy_s")
all_series<-ts(data_pred2,start=c(2015,1),frequency=12)
for(i in (96-1):(sample-1-1)){                           #lag
arimax<-auto.arima(all_series[1:i,1],seasonal = TRUE, xreg=all_series[1:i,c('Energy_i','Energy_s','Oil_p')])
new<-matrix(all_series[(i+1),c('Energy_i','Energy_s','Oil_p')],ncol=3)
colnames(new)<-c('Energy_i','Energy_s','Oil_p')
pred_arimax<-forecast::forecast(arimax,xreg=new)
predictions_uni[(i-(96-1)+1),2]<-pred_arimax$mean
lower_uni[(i-(96-1)+1),2]<-pred_arimax$lower[,2]
up_uni[(i-(96-1)+1),2]<-pred_arimax$upper[,2]
}
```

```{r}
#Arimax with text data
data_pred2<-cbind(data_pred$Renixx,lag(data_pred[,2:4]))[2:length(data_pred[,1]),]
rownames(data_pred2)<-NULL
colnames(data_pred2)<-c('Renixx','Energy_i','Oil_p',"Energy_s")
all_series<-ts(data_pred2,start=c(2015,1),frequency=12)
for(i in (96-1):(sample-1-1)){                           #lag
arimax<-auto.arima(all_series[1:i,1],seasonal = TRUE, xreg=all_series[1:i,c('Energy_i','Energy_s')])
new<-matrix(all_series[(i+1),c('Energy_i','Energy_s')],ncol=2)
colnames(new)<-c('Energy_i','Energy_s')
pred_arimax<-forecast::forecast(arimax,xreg=new)
predictions_uni[(i-(96-1)+1),3]<-pred_arimax$mean
lower_uni[(i-(96-1)+1),3]<-pred_arimax$lower[,2]
up_uni[(i-(96-1)+1),3]<-pred_arimax$upper[,2]
}
```


```{r}
#VECM
for(i in 96:(sample-1)){
all_series<-ts(data_pred[1:i,c('Renixx','Energy_i','Oil_p',"Energy_s")],start=c(2015,1),frequency=12)
var <- VARselect(all_series, type = "both", lag.max = 12, season = 12)
p<-var$selection["FPE(n)"]

var_all = VAR(all_series,p, season = 12,type = 'both') #3 for cointegration

res<-matrix(c(var_all[["varresult"]][["Renixx"]][["residuals"]],var_all[["varresult"]][["Energy_i"]][["residuals"]],var_all[["varresult"]][["Energy_s"]][["residuals"]],var_all[["varresult"]][["Oil_p"]][["residuals"]]), ncol = 4)
mvshapiro_test(res)
vars::arch.test(var_all,multivariate.only = TRUE)   
normality.test(var_all,multivariate.only = TRUE)  
serial.test(var_all, type = "PT.asymptotic",lags.pt = 12)
#plot(stability(var_all), nc = 2)

nlags=max((p-1),2)
jotest_all=ca.jo(all_series, type="eigen", K=nlags, ecdet="trend", spec= "transitory", season = 12)
summary(jotest_all)
jotest_all=ca.jo(all_series, type="trace", K=nlags, ecdet="trend", spec= "transitory", season = 12)
summary(jotest_all)

lttest(jotest_all, r=sum(jotest_all@teststat>jotest_all@cval[,2]))

y.VEC <- cajorls(jotest_all, r=sum(jotest_all@teststat>jotest_all@cval[,2]))
y.VEC
summary(y.VEC$rlm)

#urca::plotres(jotest_all)
v0.VAR <- vec2var(jotest_all, r = sum(jotest_all@teststat>jotest_all@cval[,2]))
v0.VAR

impresp <- irf(v0.VAR)
#plot(impresp)

fevd(v0.VAR)$Renixx

pred_var<-predict(v0.VAR, n.ahead = 1)
predictions_uni[(i-96+1),4]<-pred_var$fcst$Renixx[,1]
lower_uni[(i-96+1),4]<-pred_var$fcst$Renixx[,2]
up_uni[(i-96+1),4]<-pred_var$fcst$Renixx[,3]
}
```


```{r}
##BSTS
for(i in (96-1):(sample-1-1)){ #lag
(ss<- list())
ss <- AddAutoAr(ss, data_pred2[1:i,c('Renixx')])
ss <- AddLocalLevel(ss,data_pred2[1:i,c('Renixx')])
ss <- AddSeasonal(ss,data_pred2[1:i,c('Renixx')], nseasons=12)
bsts.model2 <- bsts(Renixx ~ .,
               state.specification = ss,
               niter = 1000,
               data = data_pred2[1:i,],
               ping=0,
               seed=2016,
               expected.model.size = 3)
burn <- SuggestBurn(0.1, bsts.model2)
pred_bayesian2 <- predict(bsts.model2,data_pred2[(i+1),], horizon = 1,burn = burn, quantiles = c(.025, .975))
##Saving
predictions_uni[(i-(96-1)+1),5]<-pred_bayesian2$mean
lower_uni[(i-(96-1)+1),5]<-pred_bayesian2$interval[1,]
up_uni[(i-(96-1)+1),5]<-pred_bayesian2$interval[2,]
}
```

```{r}
#ANN
all_series<-ts(data_pred2,start=c(2015,1),frequency=12)
for(i in (96-1):(sample-1-1)){ #lag
ann_rsme<-c()
pred_ann_c<-c()
best_params_ann = 0
best_score = 0
count<-
j<-round(i*0.8)
count<-1
#Validation
for(dec in c(0,0.1,0.2,0.3,0.4,0.5)){
  for(rep in c(20)){
Ann<-nnetar(all_series[1:j,1],frequency = 12,decay=dec,repeats = rep)
pred_ann_c <- as.numeric(forecast::forecast(Ann, PI=TRUE,h=length((j+1):i))$mean)
ann_rsme <- Metrics::rmse(all_series[((j+1):i),c('Renixx')], pred_ann_c)
if(count == 1){
            best_params_ann = c(dec,rep)
            best_score = ann_rsme
            count = count + 1
            }
else if(ann_rsme < best_score){
            best_params_ann = c(dec,rep)
            best_score = ann_rsme
        }
}
}
#Final model
Ann<-nnetar(ts(all_series[1:i,1],frequency = 1),repeats = best_params_ann[2],decay=best_params_ann[1],lambda='auto')
pred_ann<-forecast::forecast(Ann, PI=TRUE,h=1)
predictions_uni[(i-(96-1)+1),6]<-pred_ann$mean
lower_uni[(i-(96-1)+1),6]<-pred_ann$lower[,2]
up_uni[(i-(96-1)+1),6]<-pred_ann$upper[,2]
print(i)
}
```


```{r}
Metrics::smape(predictions_uni[,1],data_pred0$Renixx)
Metrics::smape(predictions_uni[,2],data_pred0$Renixx)
Metrics::smape(predictions_uni[,3],data_pred0$Renixx)
Metrics::smape(predictions_uni[,4],data_pred0$Renixx)
Metrics::smape(predictions_uni[,5],data_pred0$Renixx)
Metrics::smape(predictions_uni[,6],data_pred0$Renixx)

Metrics::rmse(predictions_uni[,1],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_uni[,2],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_uni[,3],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_uni[,4],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_uni[,5],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_uni[,6],data_pred0$Renixx)/sd(data_pred0$Renixx)*100

sum(lower_uni[,1]<data_pred0$Renixx&data_pred0$Renixx<up_uni[,1])
sum(lower_uni[,2]<data_pred0$Renixx&data_pred0$Renixx<up_uni[,2])
sum(lower_uni[,3]<data_pred0$Renixx&data_pred0$Renixx<up_uni[,3])
sum(lower_uni[,4]<data_pred0$Renixx&data_pred0$Renixx<up_uni[,4])
sum(lower_uni[,5]<data_pred0$Renixx&data_pred0$Renixx<up_uni[,5])
sum(lower_uni[,6]<data_pred0$Renixx&data_pred0$Renixx<up_uni[,6])

mean(up_uni[,1]-lower_uni[,1])
mean(up_uni[,2]-lower_uni[,2])
mean(up_uni[,3]-lower_uni[,3])
mean(up_uni[,4]-lower_uni[,4])
mean(up_uni[,5]-lower_uni[,5])
mean(up_uni[,6]-lower_uni[,6])
```

```{r}
data_pred$Date<-ymd(Energy_ipred$Time[133:239])
ggplot(data_pred[65:sample,], 
       aes(x =Date, 
           y = Renixx),
) +
  geom_line(aes(y=c(data_pred$Renixx[65:96],predictions_uni[,3]),color='Prediction'))+ 
geom_ribbon(aes(ymin=c(data_pred$Renixx[65:96],lower_uni[,3]), ymax=c(data_pred$Renixx[65:96],up_uni[,3])), alpha=0.1, fill = "green", 
              color = "green", linetype = "dotted")+
   geom_line(aes(color='Real'))+
  labs(title = 'ARIMAX,one step ahead')+xlab('Time')+ylab('Log-price')+
  scale_color_manual(name='Series',breaks=c('Real', 'Prediction'),
                     values=c('Real'='black', 'Prediction'='darkgreen'))
```

#Multi-step ahead

```{r}
Data_VECM<-as.data.frame(log(renixx_m$close))
colnames(Data_VECM)[1]<-'Renixx'

Data_VECM$Energy_i<-log(Energy_i$Absolute.Google.Search.Volume)

Data_VECM$Energy_s<-log(Energy_s$Absolute.Google.Search.Volume)

Data_VECM$Oil_p<-log(Oil_m$Price)
Data_VECM$Dummy1<-c(rep(1,36),rep(0,59),rep(0,83),rep(1,38))
Data_VECM$Dummy2<-c(rep(0,36),rep(1,59),rep(0,83),rep(0,38))

Energy_ipred <- read.csv("Predictions/Energy_spred.csv", header=TRUE, comment.char="#")
Energy_spred <- read.csv("Predictions/Energy_ipred.csv", header=TRUE, comment.char="#")
Oil_pred <- read.csv("Predictions/Oil_pred.csv", header=TRUE, comment.char="#")
Oil_pred$Date <- mdy(Oil_pred$Date)
Month <- as.Date(cut(Oil_pred$Date, "month"))
Oil_predm <- aggregate(Price ~ Month, Oil_pred, mean)
Renixx_pred <- read.csv("Predictions/Renixx_pred.csv", header=TRUE, comment.char="#")
Renixx_pred$date <- ymd(Renixx_pred$date)
Month <- as.Date(cut(Renixx_pred$date, "month"))
Renixx_predm <- aggregate(close ~ Month, Renixx_pred, mean)
data_pred0<-data.frame(log(Renixx_predm$close),log(Energy_ipred[229:239,3]),log(Oil_predm$Price),log(Energy_spred[229:239,3]),rep(0,11),rep(1,11))
colnames(data_pred0)<-c('Renixx','Energy_i','Oil_p',"Energy_s",'Dummy1','Dummy2')
data_pred<-rbind(Data_VECM[c(1:216),c('Renixx','Energy_i','Oil_p',"Energy_s",'Dummy1','Dummy2')],data_pred0)
rownames(data_pred)<-NULL
```

```{r}
iteration<-11
models<-3
sample<-length(data_pred[,1])
predictions_multi<-as.data.frame(matrix(nrow=iteration,ncol=models))
colnames(predictions_multi)<-c('ARIMA','VAR','NNETAR')
up_multi<-as.data.frame(matrix(nrow=iteration,ncol=models))
colnames(up_multi)<-c('ARIMA','VAR','NNETAR')
lower_multi<-as.data.frame(matrix(nrow=iteration,ncol=models))
colnames(lower_multi)<-c('ARIMA','VAR','NNETAR')
```

```{r}
#Arima
all_series<-ts(data_pred,start=c(2005,1),frequency=12)
arima<-auto.arima(all_series[1:216,c('Renixx')],seasonal = TRUE)
pred_arima<-forecast::forecast(arima,h=11)
predictions_multi[,1]<-pred_arima$mean
lower_multi[,1]<-pred_arima$lower[,2]
up_multi[,1]<-pred_arima$upper[,2]
```

```{r}
#VECM
all_series<-ts(data_pred[1:216,c('Renixx','Energy_i','Oil_p',"Energy_s")],start=c(2005,1),frequency=12)
var <- VARselect(all_series, type = "both", lag.max = 12, season = 12)
p<-var$selection["FPE(n)"]
exo<-as.matrix(data_pred[1:216,c('Dummy1','Dummy2')],ncol=2,nrow=11)
colnames(exo)<-c('exo1','exo2')

var_all = VAR(all_series,p, season = 12,type = 'both',exogen=exo) #3 for cointegration

res<-matrix(c(var_all[["varresult"]][["Renixx"]][["residuals"]],var_all[["varresult"]][["Energy_i"]][["residuals"]],var_all[["varresult"]][["Energy_s"]][["residuals"]],var_all[["varresult"]][["Oil_p"]][["residuals"]]), ncol = 4)
mvshapiro_test(res)
vars::arch.test(var_all,multivariate.only = TRUE)   
normality.test(var_all,multivariate.only = TRUE)  
serial.test(var_all, type = "PT.asymptotic",lags.pt = 12)

nlags=max(3,p)-1
jotest_all=ca.jo(all_series, type="eigen", K=nlags, ecdet="trend", spec= "transitory", season = 12,dumvar=exo)
summary(jotest_all)
jotest_all=ca.jo(all_series, type="trace", K=nlags, ecdet="trend", spec= "transitory", season = 12,dumvar=exo)
summary(jotest_all)

lttest(jotest_all, r=sum(jotest_all@teststat>jotest_all@cval[,2]))

y.VEC <- cajorls(jotest_all, r=sum(jotest_all@teststat>jotest_all@cval[,2]))
y.VEC
summary(y.VEC$rlm)

v0.VAR <- vec2var(jotest_all, r = sum(jotest_all@teststat>jotest_all@cval[,2]))
v0.VAR

exo<-matrix(c(rep(0,11),rep(1,11)),ncol=2,nrow=11)
colnames(exo)<-c('exo1','exo2')
pred_var<-predict(v0.VAR, n.ahead = 11, dumvar=exo)
predictions_multi[,2]<-pred_var$fcst$Renixx[,1]
lower_multi[,2]<-pred_var$fcst$Renixx[,2]
up_multi[,2]<-pred_var$fcst$Renixx[,3]
```

```{r}
#ANN
data_pred2<-cbind(data_pred$Renixx,lag(data_pred[,2:4]))[2:length(data_pred[,1]),]
rownames(data_pred2)<-NULL
colnames(data_pred2)<-c('Renixx','Energy_i','Oil_p',"Energy_s")
all_series<-ts(data_pred2[1:216,],start=c(2005,2),frequency=12)
ann_rsme<-c()
pred_ann_c<-c()
best_params_ann = 0
best_score = 0
j<-round(216*0.8)
count<-1
#Validation
for(dec in c(0,0.1,0.2,0.3,0.4,0.5)){
  for(rep in c(20)){
Ann<-nnetar(all_series[1:j,1],frequency = 12,decay=dec,repeats = rep)
pred_ann_c <- as.numeric(forecast::forecast(Ann, PI=TRUE,h=length((j+1):216))$mean)
ann_rsme <- Metrics::rmse(all_series[((j+1):216),c('Renixx')], pred_ann_c)
if(count == 1){
            best_params_ann = c(dec,rep)
            best_score = ann_rsme
            count = count + 1
            }
else if(ann_rsme < best_score){
            best_params_ann = c(dec,rep)
            best_score = ann_rsme
        }
}
}
#Final model
Ann<-nnetar(ts(all_series[1:216,1],frequency = 1),repeats = best_params_ann[2],decay=best_params_ann[1],lambda='auto')
pred_ann<-forecast::forecast(Ann, PI=TRUE, h=11)
predictions_multi[,3]<-pred_ann$mean
lower_multi[,3]<-pred_ann$lower[,2]
up_multi[,3]<-pred_ann$upper[,2]
```


```{r}
Metrics::smape(predictions_multi[,1],data_pred0$Renixx)
Metrics::smape(predictions_multi[,2],data_pred0$Renixx)
Metrics::smape(predictions_multi[,3],data_pred0$Renixx)

Metrics::rmse(predictions_multi[,1],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_multi[,2],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
Metrics::rmse(predictions_multi[,3],data_pred0$Renixx)/sd(data_pred0$Renixx)*100
```

```{r}
data_pred$Date<-ymd(Energy_ipred$Time[13:239])
ggplot(data_pred[185:sample,], 
       aes(x =Date, 
           y = Renixx),
) +
  geom_line(aes(y=c(data_pred$Renixx[185:216],predictions_multi[,2]),color='Prediction'))+ 
geom_ribbon(aes(ymin=c(data_pred$Renixx[185:216],lower_multi[,2]), ymax=c(data_pred$Renixx[185:216],up_multi[,2])), alpha=0.1, fill = "green", 
              color = "green", linetype = "dotted")+
   geom_line(aes(color='Real'))+
  labs(title = 'VECM, multi-step ahead')+xlab('Time')+ylab('Log-price')+
  scale_color_manual(name='Series',breaks=c('Real', 'Prediction'),
                     values=c('Real'='black', 'Prediction'='darkgreen'))
```

#Finance implications

```{r}
all_series<-ts(Data_3[c(1:216),c('Renixx','OFR','Oil_p')],start=c(2005,1), end=c(2022,12),frequency=12)

var <- VARselect(all_series, type = "none", lag.max = 12, season = 12)
var$selection["FPE(n)"]

var_OFR <- VAR(all_series,1, season = 12,type = 'none')
summary(var_OFR)
causality(var_OFR, cause = "OFR")
causality(var_OFR, cause = "Renixx")
causality(var_OFR, cause = "Oil_p")
xtable(var_OFR$varresult$OFR)
```

##VAR 

```{r}
vars::arch.test(var_OFR,multivariate.only = TRUE)   
normality.test(var_OFR,multivariate.only = TRUE)  
serial.test(var_OFR, type = "BG",lags.bg = 12)
plot(stability(var_OFR), nc = 2)

impresp <- irf(var_OFR,impulse = 'Renixx', response = 'OFR')
impresp$irf$Renixx <- -impresp$irf$Renixx #negative shock
impresp$Lower$Renixx <- -impresp$Lower$Renixx
impresp$Upper$Renixx <- -impresp$Upper$Renixx
plot(impresp, main='Impulse Response Function from RENIXX, negative shock',ylab='FSI')

plot(fevd(var_OFR),nr = 2)
fevd(var_OFR)
```
